A scenario for a Planckian standard kilogram

### 2017-10-31 manuscript_gravitationalconstant.tex

\textbf{Radiation pressure:} The envelope of the cavity needs to protect the two body-system from the constant bombardment of radiation and charged particles from the Sun. Since the activity of the Sun varies much, the spacecraft must be able to correct its position around the freely falling masses inside the cavity.\\
Furthermore, there is thermal radiation inside the cavity, that could disturb the orbit.
Using the Stefan-Boltzmann law, we can relate radiation pressure to the temperature inside the cavity. If we then specify to balls of diameter $d$, we can express the force on the balls as 
\begin{equation}
F = \frac{2 \pi^6 d^2 {k_\mathrm{B}}^4 T^4}{45 c^3 h^3}.
\end{equation}
As we can see, this force is highly temperature dependent. In order to minimize this influence, one should deploy a temperature as small as possible in the cavity. In fact, if a cooling system based on liquid helium is installed, temperature can be brought down to about 4 K. Such a low temperature results in a force on the balls in the order of $10^{-16}$ N. Compared to the gravitational force, this is orders of magnitude less and thus we don't have to worry about effects of radiation pressure.\\

\textbf{Influence of the position measurement:} 
The photons of the laser that measures the position of the two balls exert an additional force. If a laser with a power of 1 mW is used and not more than five measurements every second are performed, the force on the balls is in the same order as the influence of thermal radiation and thus negligible too.\\

@@

In this section we will show how the gravitational constant can be
recovered from the position data of the two balls that the laser
interferometry measurement would provide.  But first we have to find
the orbital equation of motion for the two balls.
\subsection{Orbital equation of motion}
From Newtonian dynamics the equation of motion for a clean two-body system is given by 
\begin{equation}
\label{eq:eomeasy}
\ddot{\vec{r}} = - \frac{G M\, \vec{r}}{r^3}.
\end{equation}
Here $M$ denotes the sum of the masses $m_1$ and $m_2$, $G$ is the
gravitational constant and $r$ is the relative distance between the
two bodies. For simplicity we will abbreviate the product of $GM$ with
$k$.  Since our two-body system is in a rotating frame, we have to add
the term developed before for the Coriolis acceleration, as well as a
correction term for centrifugal and tidal forces. The latter can be
written with a constant symmetric $3 \times 3$ matrix $C$ as follows
\begin{equation}
\label{eq:tidal}
\ddot{\vec{r}}_{\mathrm{C}} = C \cdot \vec{r},
\end{equation}
or in terms of the individual components 
\begin{equation}
\label{eq:tidacom}
\left(\ddot{r}_\mathrm{C}\right)_i = \sum_{k=1}^{3} C_{ik} \: r_k.
\end{equation}
The reference frame is chosen such that the barycenter of the two-body
system is freely falling and has no tidal or fictitious forces acting
on it.  The balls, however, do experience such forces, because they
are not at the centre of mass.  The $C$ term can be thought of as a
first-order expansion of these forces.  So the full equation of motion
for the two balls looks as follows:
\begin{equation}
\label{eq:eom}
\ddot{\vec{r}} = - \frac{k\,\vec{r}}{r^3} - 2 \; \vec{\omega} \times \dot{\vec{r}} + C \cdot \vec{r}
\end{equation}
These are actually three coupled equations of motion, one for each coordinate axis. Since all are second order differential equations, this leaves us with six initial conditions. In other words to determine the orbit, we need to give six starting values.

We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the Earth-Sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.

@@

The main contribution to the tidal term goes with $\omega^2$ and lies in the orbital plane. We are free to choose the $x$ and $y$ axis, so we choose it such that it is in $x$-direction. The other five components then should be smaller, but we give some extra range to account for possible other perturbations.
The initial values for the positions can be guessed very accurately, since they are directly measured precisely. The initial velocity can be calculated easily and matched to the initial positions, therefore also yielding narrow parameter ranges.

>>>>

\textbf{Radiation pressure:} The envelope of the cavity needs to protect the two body-system from the constant bombardment of radiation and charged particles from the Sun. Since the activity of the Sun varies much, the spacecraft must be able to correct its position around the freely falling masses inside the cavity.\\
Furthermore, there is thermal radiation inside the cavity, that could disturb the orbit.
Using the Stefan-Boltzmann law, we can relate radiation pressure to the temperature inside the cavity. If we then specify to balls of diameter $d$, we can express the force on the balls as 
\begin{equation}
F = \frac{2 \pi^6 d^2 {k_\mathrm{B}}^4 T^4}{45 c^3 h^3}.
\end{equation}
As we can see, this force is highly temperature dependent. In order to minimize this influence, one should deploy a temperature as small as possible in the cavity. In fact, if a cooling system based on liquid helium is installed, temperature can be brought down to about 4 K. Such a low temperature results in a force on the balls in the order of $10^{-16}$ N. Compared to the gravitational force, this is orders of magnitude less and thus we don't have to worry about effects of radiation pressure.

\textbf{Influence of the position measurement:} 
The photons of the laser that measures the position of the two balls exert an additional force. If a laser with a power of 1 mW is used and not more than five measurements every second are performed, the force on the balls is in the same order as the influence of thermal radiation and thus negligible too.

@@

In this section we will show how the gravitational constant can be
recovered from the position data of the two balls that the laser
interferometry measurement would provide.  But first we have to find
the orbital equation of motion for the two balls.

\subsection{Orbital equation of motion}
From Newtonian dynamics the equation of motion for a clean two-body system is given by 
\begin{equation}
\label{eq:eomeasy}
\ddot{\vec{r}} = - \frac{G M\, \vec{r}}{r^3}.
\end{equation}
Here $M$ denotes the sum of the masses $m_1$ and $m_2$, $G$ is the
gravitational constant and $r$ is the relative distance between the
two bodies. For simplicity we will abbreviate the product of $GM$ with
$k$.  Since our two-body system is in a rotating frame, we have to add
the term developed before for the Coriolis acceleration, as well as a
correction term for centrifugal and tidal forces. The latter can be
written with a constant symmetric $3 \times 3$ matrix $C$ as follows
\begin{equation}
\label{eq:tidal}
\ddot{\vec{r}}_{\mathrm{C}} = C \cdot \vec{r},
\end{equation}
or in terms of the individual components 
\begin{equation}
\label{eq:tidacom}
\left(\ddot{r}_\mathrm{C}\right)_i = \sum_{k=1}^{3} C_{ik} \: r_k.
\end{equation}
The reference frame is chosen such that the barycenter of the two-body
system is freely falling and has no tidal or fictitious forces acting
on it.  The balls, however, do experience such forces, because they
are not at the centre of mass.  The $C$ term can be thought of as a
first-order expansion of these forces.  So the full equation of motion
for the two balls looks as follows:
\begin{equation}
\label{eq:eom}
\ddot{\vec{r}} = - \frac{k\,\vec{r}}{r^3} - 2 \; \vec{\omega} \times \dot{\vec{r}} + C \cdot \vec{r}
\end{equation}

These are actually three coupled equations of motion, one for each spatial coordinate. Since all are second order differential equations, we need to give six initial conditions to determine the orbit.\\
We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the Earth-Sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.
We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the Earth-Sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.

@@

The main contribution to the tidal term goes with $\omega^2$ and lies in the orbital plane. We are free to choose the $x$ and $y$ axis, so we choose it such that the $\omega^2$-contribution is in $x$-direction. The other five components then should be smaller, but we give some extra range to account for possible other perturbations.
The initial values for the positions can be guessed very accurately, since they are directly measured precisely. The initial velocity can be calculated easily and matched to the initial positions, therefore also yielding narrow parameter ranges.
